---
title: "Full Bee Sim"
output: html_notebook
---

```{r Packages}
library(tidyverse)
library(forcats)
library(reshape2)
library(R6)
library(av)
```

Define variables
```{r Variables}
BROOD <- 1
POLLEN <- 2
HONEY <- 3 
EMPTY <- 0

MAX_ROWS <- 20
MAX_COLS <- 20
```

Makes the beehive graph framework
https://stackoverflow.com/questions/24006361/plot-series-of-filled-hexagons-in-ggplot2
```{r Beehive Graph}

types <- c("Brood", "Pollen", "Honey", "Empty")
#hive_colors <- c("#F3DAB1","#FFF966","#DE620B","#230F11")
hive_colors <- c("#FF0000","#00FF00","#0000FF","#000000")

hex_size <- 1

hex_center_xs <- rep_len(c(seq(from = 0, by = hex_size, length.out = MAX_ROWS),
                   seq(from = hex_size/2, by = hex_size, length.out = MAX_ROWS)), MAX_ROWS*MAX_COLS)

hex_center_ys <- sort(c(rep(seq(from = 0, by = hex_size*sqrt(3), length.out=ceiling(MAX_COLS/2)), MAX_ROWS),
            rep(seq(from = hex_size*sqrt(3)/2, by = hex_size*sqrt(3), length.out=floor(MAX_COLS/2)), MAX_ROWS)),decreasing = TRUE)

hex_x = cbind(hex_center_xs + 0, hex_center_xs + 0.5, hex_center_xs + 0.5,
              hex_center_xs + 0, hex_center_xs - 0.5, hex_center_xs - 0.5) 
hex_y = cbind(hex_center_ys - 1/(sqrt(3)), hex_center_ys - 1/(2*sqrt(3)), hex_center_ys + 1/(2*sqrt(3)),
              hex_center_ys + 1/(sqrt(3)), hex_center_ys + 1/(2*sqrt(3)), hex_center_ys - 1/(2*sqrt(3)))

hexdat_x <- melt(cbind(id = 1:length(hex_center_xs), as.data.frame(hex_x)), id.vars = "id", value.name = "x")
hexdat_y <- melt(cbind(id = 1:length(hex_center_ys), as.data.frame(hex_y)), id.vars = "id", value.name = "y")

hexdat <- full_join(hexdat_y, hexdat_x)

hex_center_xs_df <- melt(cbind(id = 1:length(hex_center_xs), as.data.frame(hex_center_xs)), id.vars = "id", value.name = "x") %>% select(-variable)
hex_center_ys_df <- melt(cbind(id = 1:length(hex_center_ys), as.data.frame(hex_center_ys)), id.vars = "id", value.name = "y") %>% select(-variable)

hexdat_centers <- full_join(hex_center_xs_df, hex_center_ys_df, by = "id") 

```

This takes a given hive and turns it into a graph

```{r Hive to Graph}

hive_matrix_to_graph <- function(hive_data, bees, nurses,queen){
  hive_df <- as.data.frame(as.table(hive_data[,,1:2])) %>% pivot_wider(names_from = Var3, values_from = Freq)
  colnames(hive_df) <- c("x","y","contents","amount")
  hive_df <- hive_df %>% mutate(id = 1:(MAX_COLS*MAX_ROWS)) %>% 
                         mutate(named_content = ifelse(contents == 0, "Empty",
                                                       ifelse(contents == 1, "Brood",
                                                              ifelse(contents == 2, "Pollen",
                                                                     ifelse(contents == 3, "Honey","???")))))
  new_hexdat <- full_join(hexdat,
                          hive_df %>% select(id,named_content,amount),
                          by="id") %>% mutate(named_content = fct_relevel(as.factor(named_content),types))
  
  new_hexdat <- new_hexdat %>% mutate(amount = ifelse(named_content == "Honey",amount/MAX_HONEY,
                                                      ifelse(named_content == "Pollen",amount/MAX_POLLEN,
                                                             ifelse(named_content == "Brood",amount/MAX_BROOD,1))))
  #Adding bees to the graph
  bee_locs <- matrix(EMPTY,MAX_ROWS,MAX_COLS)
  
  for(i in 1:length(bees)){
    bee_locs[bees[[i]]$cur_X,bees[[i]]$cur_Y] <- 1
  }
  
  colnames(bee_locs) <- 1:MAX_COLS
  rownames(bee_locs) <- 1:MAX_ROWS
  
  bee_locs <- as.data.frame(as.table(bee_locs))
  
  colnames(bee_locs) <- c("col","row","is_bee")
  bee_locs <- bee_locs %>% mutate(id = 1:(MAX_COLS*MAX_ROWS))
  
  new_bee_locs <- full_join(hexdat_centers,bee_locs,by = "id") %>% filter(is_bee == 1)
  
  #Adding nurses to the graph
  nurse_locs <- matrix(EMPTY,MAX_ROWS,MAX_COLS)
  
  for(i in 1:length(nurses)){
    nurse_locs[nurses[[i]]$cur_X,nurses[[i]]$cur_Y] <- 1
  }
  
  colnames(nurse_locs) <- 1:MAX_COLS
  rownames(nurse_locs) <- 1:MAX_ROWS
  
  nurse_locs <- as.data.frame(as.table(nurse_locs))
  
  colnames(nurse_locs) <- c("col","row","is_bee")
  nurse_locs <- nurse_locs %>% mutate(id = 1:(MAX_COLS*MAX_ROWS))
  
  new_nurse_locs <- full_join(hexdat_centers,nurse_locs,by = "id") %>% filter(is_bee == 1)
  
  #Adding Queen to the Graph
  queen_locs <- matrix(EMPTY,MAX_ROWS,MAX_COLS)
  
  queen_locs[queen$cur_X,queen$cur_Y] <- 1
  
  colnames(queen_locs) <- 1:MAX_COLS
  rownames(queen_locs) <- 1:MAX_ROWS
  
  queen_locs <- as.data.frame(as.table(queen_locs))
  
  colnames(queen_locs) <- c("col","row","is_bee")
  queen_locs <- queen_locs %>% mutate(id = 1:(MAX_COLS*MAX_ROWS))
  
  new_queen_locs <- full_join(hexdat_centers,queen_locs,by = "id") %>% filter(is_bee == 1)
                            
  g <- ggplot() + 
        geom_polygon(new_hexdat, mapping = aes(x,y,group = id, fill = named_content, alpha = amount), colour = "black") +
        geom_point(new_bee_locs, mapping = aes(x,y), color = "black", size = 3) +
        geom_point(new_bee_locs, mapping = aes(x,y), color = "yellow", size = 2) +
        geom_point(new_nurse_locs, mapping = aes(x,y), color = "black", size = 3) +
        geom_point(new_nurse_locs, mapping = aes(x,y), color = "orange", size = 2) +
        geom_point(new_queen_locs, mapping = aes(x,y), color = "black", size = 3) +
        geom_point(new_queen_locs, mapping = aes(x,y), color = "purple", size = 2) +
        scale_fill_manual(values=hive_colors) +
        coord_fixed() +
        theme_classic()
  
  return(g)
}
```

```{r}
make_random_hive <- function(){
  layer1 <- sample(c(EMPTY,BROOD,HONEY,POLLEN), MAX_ROWS*MAX_COLS, replace = TRUE, prob = c(0.1,0,1,1))
  
  #Second two rows are for brood honey and pollen
  rand_hive <- array(c(layer1,rep(EMPTY,MAX_ROWS*MAX_COLS),
                       rep(0,MAX_ROWS*MAX_COLS),
                       rep(0,MAX_ROWS*MAX_COLS)), dim = c(MAX_ROWS,MAX_COLS, 4))
  
  for(x in 1:MAX_ROWS){
    for(y in 1:MAX_COLS){
      layer2_val <- ifelse(rand_hive[x,y,1] == BROOD, sample(1:MAX_BROOD, 1),
                          ifelse(rand_hive[x,y,1] == HONEY, sample(1:MAX_HONEY,1),
                                 ifelse(rand_hive[x,y,1] == POLLEN, sample(1:MAX_POLLEN,1),
                                        0)))
      rand_hive[x,y,2] <- layer2_val
    }
  }
  
  return(rand_hive)
}
```


Moving through hexagons (https://www.redblobgames.com/grids/hexagons/)
If we store the hexagon values in a 2D grid we need to know which "squares" are actually next to each other on a hexagonal grid

On a square grid we can move from [X,Y] to [X+1,Y], [X-1,Y], [X,Y+1], and [X,Y-1], if we remove diagonals

On a hexagonal grid (using what they call “odd-r” horizontal layout) we can move from [X,Y] to:
    [X-1,Y-1], [X-1,Y], [X-1,Y+1], [X,Y+1], [X+1,Y], [X,Y-1]
    
But that's *just* for the odd rows (1 indexed). For the even rows it's: 
    [X,Y-1], [X-1,Y], [X,Y+1], [X+1,Y+1], [X+1,Y], [X+1,Y-1]

```{r Hex Movement}

hex_move <- function(cur_X,cur_Y,prev_X,prev_Y){
  even_row_legal_hexes <- tibble(
                                x = c(cur_X,cur_X-1,cur_X,cur_X+1,cur_X+1,cur_X+1),
                                y = c(cur_Y-1,cur_Y,cur_Y+1,cur_Y+1,cur_Y,cur_Y-1)
                               )
  odd_row_legal_hexes <- tibble(
                                x = c(cur_X-1,cur_X-1,cur_X-1,cur_X,cur_X+1,cur_X),
                                y = c(cur_Y-1,cur_Y,cur_Y+1,cur_Y+1,cur_Y,cur_Y-1)
                               )
  
  #check if our row is even or odd
  if(cur_Y %% 2 == 0){
    legal_hexes <- even_row_legal_hexes
  }
  else{
    legal_hexes <- odd_row_legal_hexes
  }
  
  #Makes sure it can't move back to where it was, or to the last square it was on
  #Also R is 1 indexed, so it can't be 0 or less
  legal_hexes <- legal_hexes %>% filter(x > 0) %>% filter(y > 0) %>%
                                 filter(x <= MAX_ROWS) %>% filter(y <= MAX_COLS) %>%
                                 filter(!(x == prev_X & y == prev_Y))
  
  next_hex <- legal_hexes %>% sample_n(1)
  
  return(c(next_hex$x,next_hex$y))
}

```

For Brood Aging
```{r}

age_brood <- function(hive_data){
  #Get the x and y of all brood
  brood_xs <- data.frame(which(hive_data[,,1] == BROOD,arr.ind = TRUE))$row
  brood_ys <- data.frame(which(hive_data[,,1] == BROOD,arr.ind = TRUE))$col
  
  for(i in 1:length(brood_xs)){
    #If they aren't about to become max age add 1 to age
    if(hive_data[brood_xs[i],brood_ys[i],2] < MAX_BROOD-1){
      hive_data[brood_xs[i],brood_ys[i],2] <- hive_data[brood_xs[i],brood_ys[i],2] + 1
      hive_data[brood_xs[i],brood_ys[i],3] <- 0
      hive_data[brood_xs[i],brood_ys[i],4] <- 0
    } else{
      #If they are then empty it out
      hive_data[brood_xs[i],brood_ys[i],1] <- EMPTY
      hive_data[brood_xs[i],brood_ys[i],2] <- 0
      hive_data[brood_xs[i],brood_ys[i],3] <- 0
      hive_data[brood_xs[i],brood_ys[i],4] <- 0
    }
    
    if(hive_data[brood_xs[i],brood_ys[i],3] < BROOD_HONEY_NEEDED || 
       hive_data[brood_xs[i],brood_ys[i],4] < BROOD_POLLEN_NEEDED){
      #If they didn't get enough food then empty it out
      hive_data[brood_xs[i],brood_ys[i],1] <- EMPTY
      hive_data[brood_xs[i],brood_ys[i],2] <- 0
      hive_data[brood_xs[i],brood_ys[i],3] <- 0
      hive_data[brood_xs[i],brood_ys[i],4] <- 0
    }
    
  }
 return(hive_data)
}

```


Gets the Bee Definitions from the Agents File
```{r}
source("Bee Agent Defs.R")
```


Runs the simulation with the bees

```{r Simulation}
bees <- list()
nurses <- list()

for(n in 1:N_BEES){
  bees <- append(bees, list(Bee$new(sample(1:MAX_ROWS, 1), sample(1:MAX_COLS, 1))))
}

for(n in 1:N_NURSES){
  nurses <- append(nurses, list(Nurse$new(sample(1:MAX_ROWS, 1), sample(1:MAX_COLS, 1))))
}

queen <- Queen$new(sample(1:MAX_ROWS, 1), sample(1:MAX_COLS, 1))

#hive <- array(EMPTY, dim = c(MAX_ROWS,MAX_COLS, 2))'
hive <- make_random_hive()

#hive[5:16,5:16,1] <- BROOD
#hive[5:16,5:16,2] <- sample(1:MAX_BROOD,(16-5+1)*(16-5+1),replace = TRUE)
hive[5:16,5:7,1] <- BROOD
hive[5:16,5:7,2] <- sample(1:MAX_BROOD,(16-5+1)*(7-5+1),replace = TRUE)
hive[5:16,5:7,3] <- sample(0:BROOD_HONEY_NEEDED,(16-5+1)*(7-5+1),replace = TRUE)
hive[5:16,5:7,4] <- sample(0:BROOD_POLLEN_NEEDED,(16-5+1)*(7-5+1),replace = TRUE)
hive[5:7,5:16,1] <- BROOD
hive[5:7,5:16,2] <- sample(1:MAX_BROOD,(7-5+1)*(16-5+1),replace = TRUE)
hive[5:7,5:16,3] <- sample(0:BROOD_HONEY_NEEDED,(7-5+1)*(16-5+1),replace = TRUE)
hive[5:7,5:16,4] <- sample(0:BROOD_POLLEN_NEEDED,(7-5+1)*(16-5+1),replace = TRUE)
hive[14:16,5:16,1] <- BROOD
hive[14:16,5:16,2] <- sample(1:MAX_BROOD,(16-14+1)*(16-5+1),replace = TRUE)
hive[14:16,5:16,3] <- sample(0:BROOD_HONEY_NEEDED,(16-14+1)*(16-5+1),replace = TRUE)
hive[14:16,5:16,4] <- sample(0:BROOD_POLLEN_NEEDED,(16-14+1)*(16-5+1),replace = TRUE)
hive[5:16,14:16,1] <- BROOD
hive[5:16,14:16,2] <- sample(1:MAX_BROOD,(16-5+1)*(16-14+1),replace = TRUE)
hive[5:16,14:16,3] <- sample(0:BROOD_HONEY_NEEDED,(16-5+1)*(16-14+1),replace = TRUE)
hive[5:16,14:16,4] <- sample(0:BROOD_POLLEN_NEEDED,(16-5+1)*(16-14+1),replace = TRUE)
# hive[1,1,1] <- EMPTY
# hive[1,1,2] <- 0

hive_matrix_to_graph(hive,bees,nurses,queen)

colnames(hive) <- 1:MAX_COLS
rownames(hive) <- 1:MAX_ROWS

DAYS <- 1

for(d in 1:DAYS){
#24 hours in a day
  for(h in 1:24){
    #Bees move 6 times and hour, then they get refilled
    for(m in 1:6){
      #bees don't forage at night
      if(h <= 12){
        for(i in 1:length(bees)){
          bees[[i]]$move()
          hive <- bees[[i]]$interact_with_comb(hive)
        }
      }
      for(i in 1:length(nurses)){
        nurses[[i]]$move()
        hive <- nurses[[i]]$interact_with_comb(hive)
      }
      for(i in 1:queen$cells_per_10){
        queen$move()
        hive <- queen$interact_with_comb(hive)
      }
      # ggsave(paste0("hive_plots/hiveplot_",str_pad((d-1)*24*6+(h-1)*6+m, 5, pad = "0"),".png"),
      #        hive_matrix_to_graph(hive,bees,nurses,queen),width=7, height=7)
    }
    #Then reset bees and nurses
    for(i in 1:length(bees)){
        bees[[i]]$honey_to_deposit <- HONEY_BY_HOUR_BY_BEE
        bees[[i]]$pollen_to_deposit <- POLLEN_BY_HOUR_BY_BEE
        bees[[i]]$honey_to_eat <- EATEN_HONEY_BY_HOUR_BY_BEE
        bees[[i]]$pollen_to_eat <- EATEN_POLLEN_BY_HOUR_BY_BEE
    }
    for(i in 1:length(nurses)){
        nurses[[i]]$honey_for_brood <- BROOD_EATEN_HONEY_BY_HOUR_BY_BEE
        nurses[[i]]$pollen_for_brood <- BROOD_EATEN_POLLEN_BY_HOUR_BY_BEE
    }
    print(hive[,,1])
    print(hive[,,3:4])
  }
  hive <- age_brood(hive)
  ggsave(paste0("hive_plots/hiveplot_",str_pad((d-1)*24*6+(h-1)*6+m, 5, pad = "0"),
                "_D",str_pad(d, 3, pad = "0"),".png"),
             hive_matrix_to_graph(hive,bees,nurses,queen),width=7, height=7)
}

```

Turn images into video
```{r}
productPath <- ('/Users/Ben/Desktop/Bee-Simulation/hive_plots/')
av::av_encode_video(paste0(productPath,list.files(productPath, '*.png')), framerate = 10,
                    output = paste0("hive_videos/bee_video_brood_death_D",DAYS,"_B",N_BEES,"_N",N_NURSES,"_Q",QUEEN_CELLS_PER_HOUR,".mp4"))
```

