---
title: "Hive Parameter Analysis"
output: html_notebook
---

```{r Load Packages}
library(MASS)
library(tidyverse)
library(ggplot2)
library(readr)
library(viridis)
library(ggfortify)
library(bbmle) #For ICtab
library(car)
library(emmeans)

 
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}

```

```{r}

round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}

```


```{r Read in Data}
# trial_types <- c("none","brood","worker","queen","all")
# 
# for (t in trial_types){
#   bee_param_df <- list.files(path=paste0("hive_data/heat_",t,"/"), full.names = TRUE) %>%
#                 lapply(read_csv, show_col_types = FALSE) %>%
#                 bind_rows %>%
#                 mutate(filenames = list.files(path=paste0("hive_data/heat_",t,"/"), full.names = TRUE))
# 
#   write.csv(bee_param_df,paste0("bee_data_isaac_heat_",t,".csv"),row.names = FALSE)
# }

```

What these all mean:
n = queen cells per hour
rb = brood radius
rn = necter radius
w = total daily honey
pph = pollen ratio
ph = honey consumption ratio
pp = pollen consumption ratio
k = consumption probability value

broodMetric = average number of brood surrounding other brood
pollenRing = average min distance between honey and brood

```{r}
brood_bee_param <- read.csv("bee_data_isaac_heat_brood.csv") %>% mutate(type = "Brood") %>% mutate(wha = 0, qha = 0)
worker_bee_param <- read.csv("bee_data_isaac_heat_worker.csv") %>% mutate(type = "Worker") %>% mutate(bhd = 0, qha = 0)
queen_bee_param <- read.csv("bee_data_isaac_heat_queen.csv") %>% mutate(type = "Queen") %>% mutate(bhd = 0, wha = 0)
none_bee_param <- read.csv("bee_data_isaac_heat_none.csv") %>% mutate(type = "None") %>% mutate(bhd = 0,wha = 0, qha = 0)
all_bee_param <- read.csv("bee_data_isaac_heat_all.csv") %>% mutate(type = "All")

bee_heat_param_df <- rbind(brood_bee_param,worker_bee_param) %>% rbind(.,queen_bee_param) %>% rbind(.,all_bee_param) %>% rbind(.,none_bee_param)
```


```{r}
ggplot(bee_heat_param_df, aes(x = type, y = pBroodHeat))+
  geom_boxplot()+
  theme_classic()

heat_glm <- glm(pBroodHeat ~ 0 + type, data = bee_heat_param_df)

contrast(emmeans(heat_glm, "type"), "pairwise", adjust = "Tukey")

```


```{r}

ggplot(all_bee_param %>% na.omit(), aes(x = qha, y = bhd, color = pBroodHeat))+
  geom_point()+
  scale_color_viridis()+
  theme_classic()

rounded_all_bee_param <- all_bee_param %>% mutate(rounded_qha = round_any(qha, 0.25),
                                                 rounded_bhd = round_any(bhd, 0.25))

ggplot(rounded_all_bee_param %>% na.omit(), aes(x = rounded_qha, y = rounded_bhd, z = pBroodHeat))+
  geom_contour_filled()+
  theme_classic()


```




```{r}

ggplot(bee_heat_param_df %>% na.omit(), aes(x = broodMetric, y = pollenRing, color = pHoney))+
  geom_point(alpha = 0.1)+
  geom_rect(aes(xmin = 5, xmax = 6, ymin = 12, ymax = 30), fill = NA)+
  scale_color_viridis()+
  theme_classic()

max(bee_heat_param_df$pollenRing)

bee_heat_param_df %>% filter(broodMetric >= 5 & pollenRing >= 10)
```



Looking at correlation between heat parameter and brood in the place

```{r}
ggplot(bee_heat_param_df %>% filter(bhd > 0), aes(x = bhd, y = pBroodHeat))+
  geom_point()+
  geom_smooth(method = 'glm', formula = 'y ~ x')+
  theme_classic()

cor.test((bee_heat_param_df %>% filter(bhd > 0))$bhd,(bee_heat_param_df %>% filter(bhd > 0))$pBroodHeat)

ggplot(bee_heat_param_df %>% filter(qha > 0), aes(x = qha, y = pBroodHeat))+
  geom_point()+
  geom_smooth(method = 'glm', formula = 'y ~ x')+
  theme_classic()

cor.test((bee_heat_param_df %>% filter(qha > 0))$qha,(bee_heat_param_df %>% filter(qha > 0))$pBroodHeat)

ggplot(bee_heat_param_df %>% filter(wha > 0), aes(x = wha, y = pBroodHeat))+
  geom_point()+
  geom_smooth(method = 'glm', formula = 'y ~ x')+
  theme_classic()

cor.test((bee_heat_param_df %>% filter(wha > 0))$wha,(bee_heat_param_df %>% filter(wha > 0))$pBroodHeat)
```

Let's make some stepwise models! Just with the all data for now so there aren't so many zeroes

```{r}

hive_pca <- prcomp(bee_heat_param_df %>% select(-c(...1, trial_n, type, days, n,
                                                   pBrood, pPollen, pHoney, pEmpty,
                                                   pBroodHeat, pPollenHeat, pHoneyHeat, pEmptyHeat,
                                                   broodMetric,pollenRing)),
                   center = TRUE,scale. = TRUE)
summary(hive_pca)
hive_pca$rotation[,1:2]

autoplot(hive_pca, colour = "pBroodHeat", loadings = TRUE, loadings.label = TRUE,
         data = bee_heat_param_df)+
  scale_color_viridis()+
  theme_classic()

ggplot(bee_heat_param_df, aes(pPollen, pBroodHeat))+
  geom_point()+
  theme_classic()

ggplot(bee_heat_param_df, aes(wha, pPollen))+
  geom_point()+
  theme_classic()

ggplot(bee_heat_param_df, aes(pph, pp, color = pPollen))+
  geom_point()+
  scale_color_viridis()+
  theme_classic()
```

```{r}
heat_brood_glm <- glm(pBroodHeat ~ ., data = all_bee_param %>% dplyr::select(-c(...1, trial_n, type, days, n,
                                                   pBrood, pPollen, pHoney, pEmpty,
                                                   pPollenHeat, pHoneyHeat, pEmptyHeat,
                                                   broodMetric,pollenRing)))

stepped_model <- stepAIC(heat_brood_glm, direction = "both")


Anova(stepped_model)

ggplot(all_bee_param, aes(bhd, pBroodHeat, color = pBroodHeat))+
  geom_point()+
  scale_color_viridis()+
  theme_classic()

ggplot(all_bee_param, aes(qha, pBroodHeat, color = pBroodHeat))+
  geom_point()+
  scale_color_viridis()+
  theme_classic()
```

now let's see about the order of those models

```{r}

bhd_m <- glm(pBroodHeat ~ bhd, data = all_bee_param)
bhd_2_m <- glm(pBroodHeat ~ poly(bhd,2), data = all_bee_param)
bhd_3_m <- glm(pBroodHeat ~ poly(bhd,3), data = all_bee_param)
bhd_ln_m <- glm(pBroodHeat ~ log(bhd), data = all_bee_param)

ICtab(bhd_m,bhd_2_m,bhd_3_m,bhd_ln_m)

ggplot(all_bee_param, aes(bhd, pBroodHeat, color = pBroodHeat))+
  geom_point()+
  geom_smooth(method = "glm", formula = "y ~ poly(x,2)", se = FALSE) +
  scale_color_viridis()+
  theme_classic()


qha_m <- glm(pBroodHeat ~ qha, data = all_bee_param)
qha_2_m <- glm(pBroodHeat ~ poly(qha,2), data = all_bee_param)
qha_3_m <- glm(pBroodHeat ~ poly(qha,3), data = all_bee_param)
qha_ln_m <- glm(pBroodHeat ~ log(qha), data = all_bee_param)

ICtab(qha_m,qha_2_m,qha_3_m,qha_ln_m)

ggplot(all_bee_param, aes(qha, pBroodHeat, color = pBroodHeat))+
  geom_point()+
  geom_smooth(method = "glm", formula = "y ~ x", se = FALSE) +
  scale_color_viridis()+
  theme_classic()

```

From the none data, looking at the good ones. The pollen ring metric seems a bit off
```{r}

good_hives <- c("hiveplot_00720_D030_T00111_none.pdf","hiveplot_00720_D030_T00125_none.pdf","hiveplot_00720_D030_T00211_none.pdf",
                "hiveplot_00720_D030_T00219_none.pdf","hiveplot_00720_D030_T00333_none.pdf","hiveplot_00720_D030_T00530_none.pdf",
                "hiveplot_00720_D030_T00619_none.pdf","hiveplot_00720_D030_T00706_none.pdf","hiveplot_00720_D030_T00923_none.pdf",
                "hiveplot_00720_D030_T00928_none.pdf","hiveplot_00720_D030_T01430_none.pdf","hiveplot_00720_D030_T01430_none.pdf",
                "hiveplot_00720_D030_T01720_none.pdf","hiveplot_00720_D030_T01914_none.pdf","hiveplot_00720_D030_T01915_none.pdf",
                "hiveplot_00720_D030_T01938_none.pdf","hiveplot_00720_D030_T02001_none.pdf","hiveplot_00720_D030_T02028_none.pdf")

good_csvs <- c()

for(hive in good_hives){
  number <- substring(hive,22,26)
  
  file <- paste0("hive_data/heat_none//Beehive Data Out Trial ",number,"_none.csv")
  
  good_csvs <- c(file,good_csvs)
}

bad_hives <- c("hiveplot_00720_D030_T00108_none.pdf","hiveplot_00720_D030_T00122_none.pdf","hiveplot_00720_D030_T00209_none.pdf",
               "hiveplot_00720_D030_T00218_none.pdf","hiveplot_00720_D030_T00236_none.pdf","hiveplot_00720_D030_T00328_none.pdf",
               "hiveplot_00720_D030_T00335_none.pdf","hiveplot_00720_D030_T00421_none.pdf","hiveplot_00720_D030_T00614_none.pdf",
               "hiveplot_00720_D030_T00635_none.pdf","hiveplot_00720_D030_T00711_none.pdf","hiveplot_00720_D030_T00935_none.pdf",
               "hiveplot_00720_D030_T01137_none.pdf","hiveplot_00720_D030_T01313_none.pdf","hiveplot_00720_D030_T01525_none.pdf",
               "hiveplot_00720_D030_T01623_none.pdf","hiveplot_00720_D030_T01725_none.pdf","hiveplot_00720_D030_T01928_none.pdf")

bad_csvs <- c()

for(hive in bad_hives){
  number <- substring(hive,22,26)
  
  file <- paste0("hive_data/heat_none//Beehive Data Out Trial ",number,"_none.csv")
  
  bad_csvs <- c(file,bad_csvs)
}
                                                                  
none_bee_param_good_bad <- none_bee_param %>% filter(filenames %in% good_csvs | filenames %in% bad_csvs) %>%
                                          mutate(bool_good_or_bad = ifelse(filenames %in% good_csvs,1,0),
                                                 good_or_bad = ifelse(filenames %in% good_csvs,"Good","Bad"))
none_bee_param_good_bad
```
Step AIC to see what predicts good or bad

```{r}

good_hive_glm <- glm(bool_good_or_bad ~ ., data = none_bee_param_good_bad %>% dplyr::select(-c(...1, trial_n, type, days, n,
                                                   pBrood, pPollen, pHoney, pEmpty,
                                                   pBroodHeat, pPollenHeat, pHoneyHeat, pEmptyHeat,
                                                   broodMetric,pollenRing)),
                     family = "binomial")

stepped_good_model <- stepAIC(good_hive_glm, direction = "both")


Anova(stepped_good_model)


```

```{r}
ggplot(none_bee_param_good_bad, aes(x = n, fill = good_or_bad))+
  geom_histogram(bins = 10)+
  theme_classic()

ggplot(none_bee_param_good_bad, aes(x = rb, fill = good_or_bad))+
  geom_histogram(bins = 10)+
  theme_classic()

ggplot(none_bee_param_good_bad, aes(x = rn, fill = good_or_bad))+
  geom_histogram(bins = 10)+
  theme_classic()

ggplot(none_bee_param_good_bad, aes(x = w, fill = good_or_bad))+
  geom_histogram(bins = 10)+
  theme_classic()

ggplot(none_bee_param_good_bad, aes(x = pph, fill = good_or_bad))+
  geom_histogram(bins = 10)+
  theme_classic()

ggplot(none_bee_param_good_bad, aes(x = ph, fill = good_or_bad))+
  geom_histogram(bins = 10)+
  theme_classic()

ggplot(none_bee_param_good_bad, aes(x = pp, fill = good_or_bad))+
  geom_histogram(bins = 10)+
  theme_classic()

ggplot(none_bee_param_good_bad, aes(x = k, fill = good_or_bad))+
  geom_histogram(bins = 10)+
  theme_classic()
```

####
OLDER CODE, looking at model type, we decided on M2
####


```{r}
rand_bee_param <- read.csv("bee_data_isaac_rand.csv") %>% mutate(type = "Rand")
m1_bee_param <- read.csv("bee_data_isaac_m1.csv") %>% mutate(type = "M1") %>% select(-trial_n)
m2_bee_param <- read.csv("bee_data_isaac_m2.csv") %>% mutate(type = "M2") %>% select(-trial_n)
m2E_bee_param <- read.csv("bee_data_isaac_m2_empty.csv") %>% mutate(type = "M2E") %>% select(-trial_n)
m2K_bee_param <- read.csv("bee_data_isaac_m2_katie.csv") %>% mutate(type = "M2K") %>% select(-trial_n)

bee_param_df <- rbind(m1_bee_param,m2_bee_param) %>% rbind(.,rand_bee_param) %>% rbind(.,m2E_bee_param) %>% rbind(.,m2K_bee_param)
```


```{r}
bee_param_df_perc <- bee_param_df %>% select(c(type,pBrood,pHoney,pPollen,pEmpty)) %>%
                                      pivot_longer(c(pBrood,pHoney,pPollen,pEmpty))
```


```{r Overall Brood and Pollen}

ggplot(bee_param_df, aes(x = broodMetric, y = pollenRing, color = type))+
  geom_point(alpha = 0.25)+
  ylim(0,16)+
  xlim(0,6)+
  theme_classic()

ggplot(bee_param_df_perc, aes(x = type, y = value, fill = name))+
  geom_boxplot()+
  theme_classic()

ggplot(m1_bee_param, aes(x = pPollen, y = pph))+
  geom_point()+
  theme_classic()

```

```{r Brood Metric GLMs}
broodMetricGLM_m0 <- glm(broodMetric ~ 1, data = m2_bee_param)
broodMetricGLM_m1 <- glm(broodMetric ~ n + rb + rn + w + pph + ph + pp + k, data = m2_bee_param)
broodMetricGLM_step <- step(broodMetricGLM_m0, direction = "both", scope = formula(broodMetricGLM_m1), trace = 0)

summary(broodMetricGLM_step)
```

```{r Pollen Ring GLMs}
pollenRingGLM_m0 <- glm(pollenRing ~ 1, data = m2_bee_param)
pollenRingGLM_m1 <- glm(pollenRing ~ n + rb + rn + w + pph + ph + pp + k, data = m2_bee_param)
pollenRingGLM_step <- step(pollenRingGLM_m0, direction = "both", scope = formula(pollenRingGLM_m1), trace = 0)

summary(pollenRingGLM_step)
```

```{r}
queen_pos_df <- read.csv("queen_pos_data_0_5.csv")

queen_pos_df_heatmap <- queen_pos_df %>% mutate(QueenXRound = round_any(QueenX,5),
                                                QueenYRound = round_any(QueenY,5)) %>%
                                        group_by(QueenXRound,QueenYRound) %>%
                                        summarise(meanAngleMoved = mean(AngleMoved),
                                                  count = n())

```


```{r}
ggplot(queen_pos_df, aes(x = QueenX, y = QueenY))+
  geom_path()+
  geom_point(aes(x=11,y=48.5),colour="red") +
  scale_color_viridis() +
  theme_classic()

ggplot(queen_pos_df, aes(x = QueenX, y = QueenY, color = Dist2Heat))+
  geom_point()+
  geom_point(aes(x=11,y=48.5),colour="red") +
  scale_color_viridis() +
  theme_classic()

ggplot(queen_pos_df, aes(x = QueenX, y = QueenY, color = cos(AngleOppHeat)))+
  geom_point()+
  geom_point(aes(x=11,y=48.5),colour="red") +
  scale_color_viridis() +
  theme_classic()

ggplot(queen_pos_df, aes(x = QueenX, y = QueenY, color = sin(AngleOppHeat)))+
  geom_point()+
  geom_point(aes(x=11,y=48.5),colour="red") +
  scale_color_viridis() +
  theme_classic()

ggplot(queen_pos_df_heatmap, aes(x = QueenXRound, y = QueenYRound, color = cos(meanAngleMoved)))+
  geom_point(size = 5)+
  geom_point(aes(x=11,y=48.5),colour="red") +
  scale_color_viridis() +
  theme_classic()

ggplot(queen_pos_df_heatmap, aes(x = QueenXRound, y = QueenYRound, color = sin(meanAngleMoved)))+
  geom_point(size = 5)+
  geom_point(aes(x=11,y=48.5),colour="red") +
  scale_color_viridis() +
  theme_classic()

ggplot(queen_pos_df_heatmap, aes(x = QueenXRound, y = QueenYRound, color = count))+
  geom_point(size = 5)+
  geom_point(aes(x=11,y=48.5),colour="red") +
  scale_color_viridis() +
  theme_classic()
```






