---
title: "OOP Bee Tester"
output: html_notebook
---

```{r}
library(R6)
library(tidyverse)
```

```{r}
BROOD <- 1
POLLEN <- 2
HONEY <- 3 
EMPTY <- 0
```


Define Starting Honeycomb

```{r}

MAX_ROWS <- 30
MAX_COLS <- 30

hive <- matrix(EMPTY,MAX_ROWS,MAX_COLS)

```

Moving through hexagons (https://www.redblobgames.com/grids/hexagons/)
If we store the hexagon values in a 2D grid we need to know which "squares" are actually next to each other on a hexagonal grid

On a square grid we can move from [X,Y] to [X+1,Y], [X-1,Y], [X,Y+1], and [X,Y-1], if we remove diagonals

On a hexagonal grid (using what they call “odd-r” horizontal layout) we can move from [X,Y] to:
    [X-1,Y-1], [X-1,Y], [X-1,Y+1], [X,Y+1], [X+1,Y], [X,Y-1]
    
But that's *just* for the odd rows (1 indexed). For the even rows it's: 
    [X,Y-1], [X-1,Y], [X,Y+1], [X+1,Y+1], [X+1,Y], [X+1,Y-1]

```{r}

hex_move <- function(cur_X,cur_Y,prev_X,prev_Y){
  even_row_legal_hexes <- tibble(
                                x = c(cur_X,cur_X-1,cur_X,cur_X+1,cur_X+1,cur_X+1),
                                y = c(cur_Y-1,cur_Y,cur_Y+1,cur_Y+1,cur_Y,cur_Y-1)
                               )
  odd_row_legal_hexes <- tibble(
                                x = c(cur_X-1,cur_X-1,cur_X-1,cur_X,cur_X+1,cur_X),
                                y = c(cur_Y-1,cur_Y,cur_Y+1,cur_Y+1,cur_Y,cur_Y-1)
                               )
  
  #check if our row is even or odd
  if(cur_Y %% 2 == 0){
    legal_hexes <- even_row_legal_hexes
  }
  else{
    legal_hexes <- odd_row_legal_hexes
  }
  
  #Makes sure it can't move back to where it was, or to the last square it was on
  #Also R is 1 indexed, so it can't be 0 or less
  legal_hexes <- legal_hexes %>% filter(x > 0) %>% filter(y > 0) %>%
                                 filter(x <= MAX_ROWS) %>% filter(y <= MAX_COLS) %>%
                                 filter(!(x == prev_X & y == prev_Y))
  
  next_hex <- legal_hexes %>% sample_n(1)
  
  return(c(next_hex$x,next_hex$y))
}

hex_move(2,2,3,3)

```

```{r}
Bee <- R6Class("Bee",
               public = list(
                 cur_X = NULL,
                 cur_Y = NULL,
                 prev_X = -1,
                 prev_Y = -1,
                 carry = NULL,
                 initialize = function(cur_X = NA, cur_Y = NA, carry = NA) {
                    self$cur_X <- cur_X
                    self$cur_Y <- cur_Y
                    self$carry <- carry
                  },
                 move = function(){
                   next_hex <- hex_move(self$cur_X,self$cur_Y,self$prev_X,self$prev_Y)
                   
                   self$prev_X <- self$cur_X
                   self$prev_Y <- self$cur_Y
                   
                   self$cur_X <- next_hex[1]
                   self$cur_Y <- next_hex[2]
                   
                   #print(c(self$cur_X,self$cur_Y))
                 },
                 alter_comb = function(current_hive){
                   current_hive[self$cur_X,self$cur_Y] <- self$carry
                   return(current_hive)
                 },
                 buzz = function(){
                   cat(paste0("Hello, I am a bee carrying ", self$carry, " at X = ", self$cur_X, " and Y = ", self$cur_Y, ".\n"))
                 }
               )
             )
```

```{r}
can_carry <- c(HONEY,POLLEN,BROOD)

n_bees = 20

bees <- list()

for(n in 1:n_bees){
  bees <- append(bees, list(Bee$new(sample(1:MAX_ROWS, 1), sample(1:MAX_COLS, 1), sample(can_carry, 1))))
}

hive <- matrix(EMPTY,MAX_ROWS,MAX_COLS)

colnames(hive) <- 1:MAX_COLS
rownames(hive) <- 1:MAX_ROWS


for(x in 1:100){
  for(i in 1:length(bees)){
    hive <- bees[[i]]$alter_comb(hive)
    bees[[i]]$move()
  }
  ggsave(paste0("hive_plots/hiveplot_",x,".png"),hive_matrix_to_graph(hive))
}

#ggsave(paste0("hive_plots/hiveplot_",x+1,".png"),hive_matrix_to_graph(hive))
```

Okay so now how do I make it work with my grapher?

```{r}

hive_matrix_to_graph <- function(hive_data){
  hive_df <- as.data.frame(as.table(hive_data))
  colnames(hive_df) <- c("x","y","contents")
  hive_df <- hive_df %>% mutate(id = 1:(MAX_COLS*MAX_ROWS)) %>% 
                         mutate(named_content = ifelse(contents == 0, "Empty",
                                                       ifelse(contents == 1, "Brood",
                                                              ifelse(contents == 2, "Pollen",
                                                                     ifelse(contents == 3, "Honey","???")))))
  new_hexdat <- full_join(hexdat,
                          hive_df %>% select(id,named_content),
                          by="id") %>% mutate(named_content = fct_relevel(as.factor(named_content),types))
  
  g <- ggplot(new_hexdat, aes(x, y)) + 
        geom_polygon(aes(group = id, fill = named_content), colour = "black") +
        scale_fill_manual(values=hive_colors) +
        coord_fixed() +
        theme_classic()
  
  return(g)
}

hive_matrix_to_graph(hive)

```

